---
title: "Breaking News Story"
author: "Ella Louise Stern"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

```{r}
## A new page featuring a sample “breaking news story” that is approximately 250 words and which features at least two pieces of information which are generated using inline R code within your Quarto document. In other words, an example would be “The election today resulted in ___ counties being won by the Republican candidate” where the __ is filled in on the fly by your inline R code. Note that to accomplish this your Quarto document will have to include the necessary R code chunks above to create the text or number objects you will insert into those spaces.Note that this is an example story you are writing - it can be based off of any dataset you choose to use, either one we’ve used previously or another you find. All elements needed to complete the code and the narrative must be part of your Quarto document, so whatever dataset you choose will need to be imported into the file at the top, etc, and then crunched appropriately to arrive at the number.To take a single value in a dataframe and convert it to a vector, you can use the pull() function if need be.
```

```{r}
library(tidyverse)
library(nycflights13) # <- make sure to install this package 
library(tidycensus)
library(janitor)
library(writexl)
library(here)
options(tigris_class = "sf")
```

```{r}
flights <- nycflights13::flights
```
```{r}
flights_leavingfromEWR <- flights %>% 
  filter(origin == "EWR") %>% 
  nrow()

flights_leavingfromEWR
```
There were `r flights_leavingfromEWR` departing EWR airport in 2013 in Newark, New Jersey. 

```{r}
longdelays_EWR <- flights
flights %>% 
  filter(origin == "EWR") %>% 
  filter(dep_delay < 4) %>% 
  nrow()

120835/73189

longdelays_JFK <- flights %>% 
  filter(origin == "JFK") %>% 
  filter(dep_delay < 4) %>% 
  nrow()

average_2013_flightdelay <- flights %>% 
  summarize(mean(dep_delay, na.rm = TRUE))

average_2013_flightdelay

```
On average, flights in 2013 were delayed by `r average_2013_flightdelay` minutes.

```{r}
average_delays_EWR <- flights %>% 
  filter(origin == "EWR") %>% 
  summarize(mean(dep_delay, na.rm = TRUE))

average_delays_EWR
```

BREAKING NEWS STORY

On average, flights departing Newark were delayed by `r average_delays_EWR` minutes.

In 2013, air travel faced significant delays, with flights departing from Newark Airport experiencing longer wait times than the average flight. According to recent data, `r flights_leavingfromEWR` flights departed from Newark Airport in that year, with an average delay of `r average_delays_EWR` minutes.

This figure is notably higher than the average delay experienced by all flights in 2013, which was `r average_2013_flightdelay` minutes. The discrepancy between the two figures suggests that Newark Airport faced unique challenges that impacted its ability to operate flights efficiently.

The reasons behind the delays are varied, ranging from adverse weather conditions to mechanical issues and operational challenges. However, the airport and airlines have taken steps to address the problem, including investing in new technologies, streamlining processes, and improving communication between airlines and air traffic control.

Despite the challenges faced, air travel in 2013 saw continued growth, with more people taking to the skies than ever before. This growth is expected to continue, with more people seeking the convenience and speed of air travel in the years to come.

As air travel continues to evolve, it is essential for airports and airlines to work together to minimize delays and ensure that passengers can travel safely and efficiently. With ongoing improvements in technology and operations, the future of air travel looks bright.
