---
title: "Analysis Walkthrough Example"
author: "Ella Stern"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---

We'll start by loading. our libraries.

```{r}

#load libraries
library(tidyverse)
library(lubridate)

#load data
payments <- readRDS("trump_campaign_payments.rds")


```

<br>

How much was spent in the 2017 vs. 2018 reporting year?

```{r}

payments %>% 
  group_by(disbursement_year) %>% 
  summarise(totalspent = sum(disbursement_amount)) 

## since there is only 2017 and 2018 data i grouped by the code for disbursement year and found the sum total for amount spent




```

Tally the total overall disbursement amounts for each month (distinguishing between say, Oct. 2018 v. Oct. 2019). In which month was spending the highest?

```{r}

payments %>% 
  group_by(disbursement_month, disbursement_year) %>% 
  summarise(totalspent = sum(disbursement_amount)) %>% arrange(desc(totalspent)) 

## in October 2018 the spending was the highest
## i first grouped by the month and year (being 2017 and 2018) then found the total amout spent and used the arrange(desc) command to order them highest to lowest

```

Which recipient (recipient_name) was paid the most in the 2020 election cycle?

```{r}
payments %>% 
  group_by(recipient_name) %>% 
  filter(election_cycle == 2020) %>% 
  summarise(totalrec = sum(disbursement_amount)) %>% 
  arrange(desc(totalrec)) %>% top_n(1)

# i filtered by the 2020 election cycle and summarised for the total amount received in ascending order 

```

Look at what this recipient was paid for. What service or good (see the "disbursement_description") were they paid the most for?

```{r}

payments %>% 
  filter(recipient_name == "AMERICAN MADE MEDIA CONSULTANTS, LLC") %>% 
  group_by(disbursement_description) %>% 
  summarise(total = sum(disbursement_amount)) %>% 
  arrange(desc(total))

# i filtered recipient for the recipient that recieved the most amount in 2020
# then I grouped by description and summed the disbursement amount for each description and arranged in descending order.

```

What's the most common category overall (see "disbursement_description") the campaign spent money on?

```{r}
payments %>% 
  group_by(disbursement_description) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

## the campaign spent the most common categroy was merchant fees

```

Who were the 5 biggest recipients of campaign spending on "LEGAL CONSULTING"..?

```{r}

payments %>% 
  group_by(recipient_name) %>% 
  filter(disbursement_description == "LEGAL CONSULTING") %>% 
  summarise(total = sum(disbursement_amount)) %>% 
  arrange(desc(total)) %>% top_n(5)

## i used the code from above but filtered only for this recipient. then i arranged using the function top_n which i found from the tidyverse encyclopedia online
```

Who are the top 5 individuals (vs. organization) were paid by the campaign committee?

```{r}

payments %>% 
  group_by(recipient_name) %>% 
  filter(entity_type == "IND") %>% 
  summarise(total = sum(disbursement_amount)) %>% 
  arrange(desc(total)) %>% top_n(5)

## i used the same code as above but filtered for the entity type of individuals


```

For the indvidual who was paid the most, which payment category (disbursement_description) was the top one for him or her? And how much money were they paid in that category?

\`\`\`{r}

payments %\>% group_by(disbursement_description) %\>% filter(recipient_name == "GLASSNER, MICHAEL") %\>% summarize(total = sum(disbursement_amount)) %\>% arrange(desc(total)) %\>% top_n(1)

```{r}
# i filtered for michael and found the total amount he was paid
# i then arranged for only the top amount through the command top_n(1)
```

## 
